---
title: "Auto MPG Data Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries

```{r}
library(knitr)
library(dplyr)
library(corrplot)
library(visreg)
library(ggplot2)
#library(scatterplot3d)
```
## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.


## Data Description

1. Title: Auto-Mpg Data

2. Sources:
   (a) Origin:  This dataset was taken from the StatLib library which is
                maintained at Carnegie Mellon University. The dataset was 
                used in the 1983 American Statistical Association Exposition.
   (c) Date: July 7, 1993

3. Past Usage:
    -  See 2b (above)
    -  Quinlan,R. (1993). Combining Instance-Based and Model-Based Learning.
       In Proceedings on the Tenth International Conference of Machine 
       Learning, 236-243, University of Massachusetts, Amherst. Morgan
       Kaufmann.

4. Relevant Information:

   This dataset is a slightly modified version of the dataset provided in
   the StatLib library.  In line with the use by Ross Quinlan (1993) in
   predicting the attribute "mpg", 8 of the original instances were removed 
   because they had unknown values for the "mpg" attribute.  The original 
   dataset is available in the file "auto-mpg.data-original".

   "The data concerns city-cycle fuel consumption in miles per gallon,
    to be predicted in terms of 3 multivalued discrete and 5 continuous
    attributes." (Quinlan, 1993)

5. Number of Instances: 398

6. Number of Attributes: 9 including the class attribute

7. Attribute Information:

    1. mpg:           continuous
    2. cylinders:     multi-valued discrete
    3. displacement:  continuous
    4. horsepower:    continuous
    5. weight:        continuous
    6. acceleration:  continuous
    7. model year:    multi-valued discrete
    8. origin:        multi-valued discrete
    9. car name:      string (unique for each instance)

8. Missing Attribute Values:  horsepower has 6 missing values

## Including Code

You can include R code in the document as follows:

```{r}
data <- read.table("http://mlr.cs.umass.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data",header = T, col.names = c("mpg","cylinders","displacement","horsepower","weight","acceleration","model_year", "origin","car_name"))
```

### Descriptive Analysis

```{r}
str(data)

#View(data)
```

```{r}
glimpse(data)
```

```{r}
head(data)
```

```{r}
summary(data)
```

```{r}
#factor(data$model_year)['levels']
print("Unique model years")
unique(data$model_year)

print("Unique origin")
unique(data$origin)

print("Unique cylinders")
unique(data$cylinders)
```

### Checking missing values
```{r}
anyNA(data)
is.na(data$horsepower)
```

## Data Cleaning

* Cylinders column should be factors (multi-valued discrete) not integer
```{r}
#factor(data,labels=c("I","II","III")) -- dplyr %>% method passing data$cylinder as argument to fator fn
data$cylinders = data$cylinders %>%
                 factor(labels = sort(unique(data$cylinders)))
```

* Horsepower is factor and it should be continuous numeric variable
```{r}
data$horsepower = as.numeric(levels(data$horsepower))[data$horsepower]
```
* Horsepower has some missing values. We will impute those by mean.

```{r}
#library(zoo)
#na.aggregate(DF)
#na_count <-sapply(x, function(y) sum(length(which(is.na(y)))))
#na_count <- data.frame(na_count)
#sum(is.na(data$horsepower))

colSums(is.na(data))

data$horsepower[is.na(data$horsepower)] = mean(data$horsepower,na.rm = T)
```

* Cylinders 3 & 5 has very low values. We can drop these cylinders

```{r}
#data %>% group_by(cylinders) %>% summarise(length(cylinders))

data %>% group_by(cylinders) %>% count(cylinders)

data <- data %>% filter(cylinders != 3 & cylinders != 5)

#p %>% group_by(cylinders) %>% summarise(length(cylinders))
```

* Converting Model Year to factor since it has few levels
```{r}

data$model_year = data$model_year %>%
                  factor(labels = sort(unique(data$model_year)))
```


* Converting Origin to factor since it has only 3 levels
```{r}

data$origin = data$origin %>%
                  factor(labels = sort(unique(data$origin)))
```

## Visual Analysis

[Learn ggplot](http://www.sthda.com/english/wiki/ggplot2-essentials)

* Accelration data is normaly distributed. Rest are right skewed.
```{r}
library(reshape2)

ggplot(data,aes(mpg, fill=cylinders)) +
  geom_histogram(color="black")

ggplot(data, aes(x=acceleration)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") 

ggplot(data, aes(x=horsepower)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") 

ggplot(data, aes(x=displacement)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") 

ggplot(data, aes(x=weight)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") 

d <- melt(data[,-c(8:9)])

ggplot(d,aes(value)) + 
    facet_wrap(~variable,scales = "free_x",nrow = 3) + 
    geom_histogram(colour="black", fill="red")

#ggplot(data,aes(mpg, fill = model_year)) + geom_histogram(stat = "bin")
#hist(data$mpg)
```


### Checking for outliers

[What is a Boxplot?](http://www.clayford.net/statistics/a-note-on-boxplots-in-r/)


```{r}
ggplot(data, aes(model_year,mpg,color=cylinders)) +
  geom_boxplot()
```


```{r}
ggplot(data, aes(origin,mpg)) +
  geom_boxplot()
```

* Origin 1 has heavy weighted cars (median ~ 3400)
```{r}
ggplot(data, aes(origin,weight)) +
  geom_boxplot()
```

```{r}
ggplot(data, aes(cylinders,weight,fill=cylinders)) +
  geom_boxplot()
```


```{r}
ggplot(data, aes(x=factor(cylinders),y=mpg,color=factor(cylinders)))+
  geom_boxplot(outlier.color = "red")

d <- melt(data[,-c(8:9)])

ggplot(d,aes('',value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_boxplot(outlier.colour="red", outlier.shape=16, outlier.size=2, notch=F)
```

### Scatterplot

* Miles per gallon (mpg) decreasing with increase of the weight

```{r}
ggplot(data,aes(weight,mpg)) +
  geom_point()+
  geom_smooth(method=lm)  

ggplot(data,aes(cylinders,mpg)) +
  geom_point()+
  geom_smooth(method=lm)  

ggplot(data,aes(displacement,mpg)) +
  geom_point()+
  geom_smooth(method=lm)  

ggplot(data,aes(weight, displacement)) +
  geom_point(color="red") +
  geom_smooth(method = lm)
```

* Weight, Horsepower and Displacement are highly correlated, so we can pick one attribute out of 3

```{r}
newdata <- cor(data[ , c('mpg','weight', 'displacement', 'horsepower', 'acceleration')], use='complete')
corrplot(newdata, method = "number")

```

* 6 and 8 cylinders cars are majorly built in origin 1.

```{r}
ggplot(data, aes(cylinders,fill=origin)) +
  geom_bar(position = "dodge")

ggplot(data, aes(cylinders,fill=origin)) +
  geom_bar(position = "stack")
```

* Significant drop in the car weights in origin 1. The reason behind it is increase in production of 4 cylinders cars those weighs less.

```{r}
ggplot(data, aes(model_year, y = weight, color=origin)) +
  geom_boxplot() +
  facet_wrap(~ origin) +
  xlab('Model Year') +
  ylab('Weight') +
  ggtitle('Car Weights Distributions Over Time by Region of Origin')
```

* We can see that over the year there was increase in the milege of the cars (Miles Per Gallon)

```{r}
 ggplot(data, aes(model_year,mpg,group=1))+geom_smooth()
```

* Significant drop in Car Engine's horsepower over the years
```{r}
 ggplot(data, aes(model_year,horsepower,group=1))+geom_smooth()
```

### Building Linear Model - Weight is more significant among other features and it was highly correlated to Target variable MPG

* Spliting the dataset in Train and Test (80-20)
```{r}
set.seed(100)

#80%-20% split

indexes <- sample(nrow(data), (0.80*nrow(data)), replace = FALSE)

trainData <- data[indexes, ]
testData <- data[-indexes, ]
```

* Creating the Linear Model with significant features

```{r}
model <- lm(mpg~weight+horsepower+origin+model_year+displacement+acceleration,data = data)

```

* Stats for the linear model
```{r}
summary(model)
```

* Plots for the linear model

[Plots diagnostic](http://data.library.virginia.edu/diagnostic-plots/)

```{r}
plot(model)
```

```{r}
visreg(model)
```

```{r}
predictions <- predict(model, newdata = testData)

sqrt(mean((predictions - testData$mpg)^2))
```